% !TeX root = draft-ur-v2.tex

\section{RecordIdentityBlock}

This block contains the properties related to the identity of the usage record itself.

\begin{XMLexample}
<ur:RecordIdentityBlock>
<!— Identity properties go in here -->
</ur:RecordIdentityBlock>
\end{XMLexample}



%%%%%%%%%%%%%%%



\subsection{RecordId}

A record identity uniquely defines a record in the set of all usage records of the accounting context.
\begin{itemize}
\item \emph{RecordId} MUST be present in the \emph{RecordIdentityBlock}.
\item \emph{RecordId} MUST be a string.
\end{itemize}

\begin{XMLexample}
<ur:RecordId>"host.example.org/ur/87912469269276"</ur:RecordId>
\end{XMLexample}



%%%%%%%%%%%%%%%



\subsection{CreateTime}

The time when this particular Usage Record was created.
\begin{itemize}
\item \emph{CreateTime} MUST be present in the \emph{RecordIdentityBlock}.
\item \emph{CreateTime} MUST be an ISO 8601:2004 timestamp.
\end{itemize}

\begin{XMLexample}
<ur:CreateTime>2013-05-09T09:06:52Z</ur:CreateTime>
\end{XMLexample}



%%%%%%%%%%%%%%%



\subsection{Site}

The site at which the resource is located. This property should contain a descriptive name of the group of resources which are accounted for in the record. The \emph{Site} value should be constructed in such a way that it is unique within the context where it is used.
\begin{itemize}
\item \emph{Site} MUST be a string.
\end{itemize}

\begin{XMLexample}
<ur:Site>"ACME-University"</ur:Site>
\end{XMLexample}



%%%%%%%%%%%%%%%



\subsection{Infrastructure}

The infrastructure where the resource was used (e.g., EGI or OSG).
\begin{itemize}
\item \emph{Infrastructure} MUST be a string.
\item The attribute \emph{description} SHOULD be used to give additional information on the used infrastructure.
\end{itemize}

\begin{XMLexample}
<ur:Infrastructure ur:description="U.S Open Science Grid">"OSG"</ur:Infrastructure>
\end{XMLexample}
