% !TeX root = draft-ur-v2.tex

\section{NetworkUsageBlock}

This block contains the properties related to network usage. The block may be present several times to account for different types of network.

\begin{XMLexample}
<ur:NetworkUsageBlock>
<!— Network Usage properties go in here -->
</ur:NetworkUsageBlock>
\end{XMLexample}



%%%%%%%%%%%%%%%



\subsection{NetworkClass}

The class of network used.
This is a descriptive value which allows details about the network system to be provided. The attribute \emph{NetworkResourceBandwidth} represents the maximum bandwidth allowed for this \emph{NetworkClass} expressed in bytes.

\begin{itemize}
\item \emph{NetworkClass} MUST be present in the \emph{NetworkUsageBlock}.
\item \emph{NetworkClass} MUST be a string.
%\item The value ``Ethernet'' MUST be supported.
\item The attribute \emph{NetworkResourceBandwidth} MUST be a non-zero integer.
\item This value should be defined in a community specific profile.
\end{itemize}

\begin{XMLexample}
<ur:NetworkClass ur:NetworkResourceBandwidth=100000000>"Ethernet"</ur:NetworkClass>
\end{XMLexample}



%%%%%%%%%%%%%%%



\subsection{NetworkInboundUsed}

The number of physical bytes used on the network system for inbound network traffic. The attribute \emph{SourceAddress} represents the source of the inbound network traffic.

\begin{itemize}
\item \emph{NetworkInboundUsed} MUST be present in the \emph{NetworkUsageBlock}.
\item The attribute \emph{NetworkInboundUsed} MUST be a positive integer.
\item The attribute \emph{SourceAddress} MUST be a string.
\end{itemize}

\begin{XMLexample}
<ur:NetworkInboundUsed ur:SourceAddress=192.168.1.12>14728</ur:NetworkInboundUsed> 
\end{XMLexample}



%%%%%%%%%%%%%%%



\subsection{NetworkOutboundUsed}

The number of physical bytes used on the network system for outbound network traffic. The attribute \emph{DestinationAddress} represents the destination of the outbound network traffic.

\begin{itemize}
\item \emph{NetworkOutboundUsed} MUST be present in the \emph{NetworkUsageBlock}.
\item \emph{NetworkOutboundUsed} MUST be a positive integer.
\item The attribute \emph{DestinationAddress} MUST be a string.
\end{itemize}

\begin{XMLexample}
<ur:NetworkOutboundUsed ur:DestinationAddress=192.168.1.21>14728</ur:NetworkOutboundUsed> 
\end{XMLexample}



%%%%%%%%%%%%%%%



\subsection{Charge}

Same as \emph{Charge} in \emph{ComputeUsageBlock} (see chapter \ref{ComputeCharge}).
